### 2025-11-03 Text/Thesis/Chapters/Chapter3.qmd
Versão final da tese e ficha catalográfica

The window radius $N_{ij}$ is updated dynamically and locally as follows:
$$
N_{i,j}^{\text{updated}} =
\begin{cases}
\min\{N_{ij} + 1,\, N_{\max}\}, & \text{if } C_{ij} \leq U_{ij}, \\
\max\{N_{ij} - 1,\, N_{\min}\}, & \text{if } C_{ij} > U_{ij},
\end{cases}
$$
where $N_{\max}$ and $N_{\min}$ denote the maximum and minimum allowed window size, respectively. Hence, windows *grow* in homogeneous regions (stabilizing estimates) and *shrink* in heterogeneous areas (preserving edges and avoiding class mixing).

@@

In the experiments presented in the next chapter, the Tsallis-based test statistic $S_{\widetilde{T}{\lambda}}$ will be applied to both simulated and real SAR images using adaptive windows with side lengths ranging from $5\times5$ ($N{\min}=5$) to $11\times11$ ($N_{\max}=11$), with the smoothing parameter set to $\eta = 3$. For comparison, a fixed $7\times7$ window configuration will also be evaluated.

>>>>

The window radius $N_{ij}$ is updated dynamically and locally as follows:
$$
N_{i,j}^{\text{updated}} =
\begin{cases}
\min\{N_{ij} + 1,\, N_{\max}\}, & \text{if } C_{ij} \leq U_{ij}, \\
\max\{N_{ij} - 1,\, N_{\min}\}, & \text{if } C_{ij} > U_{ij},
\end{cases}
$$
where $N_{\max}$ and $N_{\min}$ denote the maximum and minimum allowed window radius, respectively. Since $W_{ij}=2N_{ij}+1$, these bounds on the radius induce the corresponding minimum and maximum window sides $W_{\min}=2N_{\min}+1$ and $W_{\max}=2N_{\max}+1$. Hence, windows *grow* in homogeneous regions (stabilizing estimates) and *shrink* in heterogeneous areas (preserving edges and avoiding class mixing).

@@

In the experiments presented in the next chapter, the Tsallis-based test statistic $S_{\widetilde{T}{\lambda}}$ will be applied to both simulated and real SAR images using adaptive windows with side lengths ranging from $5\times5$ ($W_{\min}=5$) to $11\times11$ ($W_{\max}=11$), with the smoothing parameter set to $\eta = 3$. For comparison, a fixed $7\times7$ window configuration will also be evaluated.

### 2025-11-03 Text/Thesis/Chapters/Chapter4.qmd
Versão final da tese e ficha catalográfica

We evaluated the three test statistics $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$ on simulated images using fixed sliding windows of size $7\times7$ and, for Tsallis only, the adaptive windowing strategy with side lengths from $5\times5$ ($N_{\min}=5$) to $11\times11$ ($N_{\max}=11$) and smoothing parameter $\eta=3$. Figures&nbsp;\ref{fig:sim1t} and \ref{fig:sim2t} present, respectively for $L=5$ and $L=9$, the $p$-value maps (top row) and the corresponding binary decision maps (bottom row).

@@

We applied the three bootstrap statistics, $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$, to the five SAR intensity images in Figures&nbsp;\ref{fig:london-sar}–\ref{fig:ja-sar}. The fixed approach uses local sliding windows of size $7\times7$. The adaptive approach (used with Tsallis) varies the window side from $5\times5$ ($N_{\min}=5$) to $11\times11$ ($N_{\max}=11$), with smoothing parameter $\eta=3$, and $B=100$ bootstrap replications per window. Each figure is organized in three rows: (i) SAR scene and a matching Sentinel-2 optical image, (ii) $p$-value maps for the three tests, and (iii) binary maps at the $5\%$ level (black = heterogeneous, white = homogeneous).

@@

In addition to the visual analysis presented above, we measured execution times to assess the computational cost of the proposed methods. @tbl-exec-times reports the processing times for the three entropy-based tests (Shannon, Rényi, and Tsallis) with fixed $7\times7$ windows and, for comparison, the adaptive Tsallis variant with window sizes from $5\times5$ ($N_{\min}=5$) to $11\times11$ ($N_{\max}=11$). Times are reported in minutes and seconds (mm:ss) with $B=100$ bootstrap replications per window.

@@

\renewcommand{\arraystretch}{1.8} 
```{r}
#| label: tbl-exec-times
#| echo: false
#| message: false
#| warning: false
#| results: 'asis'
#| tbl-cap: 'Execution times for the heterogeneity detection using fixed $7\times7$ and adaptive windows ($N_{\min}=5$, $N_{\max}=11$), with $B=100$ bootstrap replications per scenario. Times are given in minutes and seconds (mm:ss). Rényi entropy uses $\lambda=0.9$ and Tsallis entropy uses $\lambda=0.85$'
#| tbl-pos: 'hbt'

>>>>

We evaluated the three test statistics $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$ on simulated images using fixed sliding windows of size $7\times7$ and, for Tsallis only, the adaptive windowing strategy with side lengths from $5\times5$ ($W_{\min}=5$) to $11\times11$ ($W_{\max}=11$) and smoothing parameter $\eta=3$. Figures&nbsp;\ref{fig:sim1t} and \ref{fig:sim2t} present, respectively for $L=5$ and $L=9$, the $p$-value maps (top row) and the corresponding binary decision maps (bottom row).

@@

We applied the three bootstrap statistics, $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$, to the five SAR intensity images in Figures&nbsp;\ref{fig:london-sar}–\ref{fig:ja-sar}. The fixed approach uses local sliding windows of size $7\times7$. The adaptive approach (used with Tsallis) varies the window side from $5\times5$ ($W_{\min}=5$) to $11\times11$ ($W_{\max}=11$), with smoothing parameter $\eta=3$, and $B=100$ bootstrap replications per window. Each figure is organized in three rows: (i) SAR scene and a matching Sentinel-2 optical image, (ii) $p$-value maps for the three tests, and (iii) binary maps at the $5\%$ level (black = heterogeneous, white = homogeneous).

@@

In addition to the visual analysis presented above, we measured execution times to assess the computational cost of the proposed methods. @tbl-exec-times reports the processing times for the three entropy-based tests (Shannon, Rényi, and Tsallis) with fixed $7\times7$ windows and, for comparison, the adaptive Tsallis variant with window sizes from $5\times5$ ($W_{\min}=5$) to $11\times11$ ($W_{\max}=11$). Times are reported in minutes and seconds (mm:ss) with $B=100$ bootstrap replications per window.

@@

\renewcommand{\arraystretch}{1.8} 
```{r}
#| label: tbl-exec-times
#| echo: false
#| message: false
#| warning: false
#| results: 'asis'
#| tbl-cap: 'Execution times for the heterogeneity detection using fixed $7\times7$ and adaptive windows ($W_{\min}=5$, $W_{\max}=11$), with $B=100$ bootstrap replications per scenario. Times are given in minutes and seconds (mm:ss). Rényi entropy uses $\lambda=0.9$ and Tsallis entropy uses $\lambda=0.85$'
#| tbl-pos: 'hbt'
